Fast Regression of the Tritium Breeding Ratio in Fusion Reactors

### 2020-06-17 paper/methodology.tex
decrease table row spacing
add extra punctuation at the end of a sentence

\begin{table}[t]
	\setlength\tabcolsep{2pt}
	\caption{\label{tbl:surrogates}Considered surrogate model families, their
		selected abbreviations and implementations. $\mathcal{H}$~denotes the
		set of hyperparameters.}
	\begin{indented}
	\item[]
		\begin{tabular}{lllr}
		\toprule
		Surrogate family & Abbr. & Impl. & $|\mathcal{H}|$ \\
		\midrule
		Support vector machines~\cite{fan2008liblinear}	& SVM & SciKit~\cite{scikit-learn} & 3 \\
		Gradient boosted trees~\cite{friedman2001greedy,friedman1999stochastic,hastie2009elements}	& GBT & SciKit & 11 \\
		Extremely randomised trees~\cite{geurts2006extremely}	& ERT & SciKit & 7 \\
		AdaBoosted decision trees$^\text{a}$~\cite{drucker1997improving}	& ABT & SciKit & 3 \\
		Gaussian process regression~\cite{williams2006gaussian}	& GPR & SciKit & 2 \\
		$k$ nearest neighbours	& KNN & SciKit & 3 \\
		Artificial neural networks	& ANN & Keras~\cite{chollet2015keras} & 2 \\
		Inverse distance weighing~\cite{shepard1968two} & IDW & SMT~\cite{SMT2019} & 1 \\
		Radial basis functions & RBF & SMT & 3 \\
		\bottomrule
		\end{tabular}\\%
		{\footnotesize $^\text{a}$Note that ABTs can be viewed as a subclass of GBTs.}
	\end{indented}
\end{table}

@@

\begin{table*}[t]
	\caption{\label{tbl:metrics}Metrics recorded in experiments. In
	formulations, we work with a training set of size $N_0$ and a test set of
size $N$, values $y^{(i)}=f(x^{(i)})$ and $\hat{y}^{(i)}=\hat{f}(x^{(i)})$
denote images of the $i$th testing sample in the MC TBR model and the surrogate
respectively. The mean $\overline{y}=N^{-1}\sum_{i=1}^N y^{(i)}$ and $P$ is the
number of input features.}
	\begin{indented}
	\item[]
		\begin{tabularx}{\textwidth}{Xrl}
		\toprule
		Regression performance metrics& Notation	& Mathematical formulation\\
		\midrule
		Mean absolute error	& MAE & $N^{-1}\sum_{i=1}^N |y^{(i)}-\hat{y}^{(i)}|$ \\
		Standard deviation of error & $S$	& $\text{StdDev}_{i=1}^N\left\{ |y^{(i)} -
		\hat{y}^{(i)}| \right\} $ \\
			Coefficient of determination & $R^2$	& $1-\sum_{i=1}^N
			\left(y^{(i)}-\hat{y}^{(i)} \right)^2\left[\sum_{i=1}^N \left(
			y^{(i)}-\overline{y} \right)^2\right]^{-1} $ \\
			Adjusted $R^2$ & $R^2_\text{adj.}$	& $1-(1-R^2)(N-1)(N-P-1)^{-1}$ \\
		\midrule
		Computational complexity metrics	& {}	& {} \\
		\midrule
		Mean training time & $\overline{t}_{\text{trn.}}$	& $(\text{wall training time of
		$\hat{f}(x)$})N_0^{-1}$  \\
			Mean prediction time & $\overline{t}_{\text{pred.}}$	& $(\text{wall prediction time of
			$\hat{f}(x)$})N^{-1}$ \\
				Relative speedup & $\omega$	& $(\text{wall evaluation$^\text{b}$ time of $f(x)$})
				(N\overline{t}_{\text{pred.}})^{-1}$ \\
		\bottomrule
		\end{tabularx}\\%
		{\footnotesize $^\text{b}$This corresponds to evaluation of the
		expensive MC TBR model on all points of the test set. In surrogates, the
		same time period is referred to as ``prediction time.''}
	\end{indented}
\end{table*}

@@

The objective of Experiment~1 is to simplify the regression task for
surrogates prone to suboptimal performance in discontinuous spaces.
To this end, training points are filtered to a single selected discrete feature
assignment, and surrogates are trained only on the remaining continuous features.
This is repeated several times to explore variances in behaviour,
particularly in 4~distinct assignments that are obtained by setting blanket and
firstwall coolant materials to one
of:~$\{\text{H\textsubscript{2}O},\isotope{He}\}$.
Experiment~2 conventionally measures surrogate performance on the full feature
space without any parameter restrictions. In both experiments, hyperparameter tuning is
facilitated by Bayesian optimisation~\cite{movckus1975bayesian}. We select the
hyperparameter configuration that produces the model that maximises $R^2$. The
process is terminated after 1000~iterations or 2~days, whichever condition is satisfied first.
The results of Experiments~1 \&~2 are depicted
in Figures~\ref{fig:exp1-time-vs-reg} \&~\ref{fig:exp2-time-vs-reg}
respectively, and described in~\Sref{sec:res-exp12}

In Experiment~3, the 20~best-performing hyperparameter configurations
per each model family are used to train surrogates on sets of various sizes to
investigate their scaling properties. In particular, we consider sets of
sizes~$\{1,2,5,10,12,15,20\}$ in thousands of samples to better characterize the
relationship of each family between sample count and the metrics of interest
listed in~\Tref{tbl:metrics}.
The results are shown in~\Fref{fig:scaling} and discussed in~\Sref{sec:res-exp3}

Finally, Experiment~4 aims
to produce surrogates suitable for practical use by retraining selected
well-scaling instances on large training sets in order to satisfy the goals of this work.
The results are displayed in~\Fref{fig:reg-performance} and
in~\Tref{tbl:exp4-detailed-results}, and summarized in~\Sref{sec:res-exp4}

>>>>

\begin{table}[t]
	\setlength\tabcolsep{2pt}
	\renewcommand{\arraystretch}{0.95}
	\caption{\label{tbl:surrogates}Considered surrogate model families, their
		selected abbreviations and implementations. $\mathcal{H}$~denotes the
		set of hyperparameters.}
	\begin{indented}
	\item[]
		\begin{tabular}{lllr}
		\toprule
		Surrogate family & Abbr. & Impl. & $|\mathcal{H}|$ \\
		\midrule
		Support vector machines~\cite{fan2008liblinear}	& SVM & SciKit~\cite{scikit-learn} & 3 \\
		Gradient boosted trees~\cite{friedman2001greedy,friedman1999stochastic,hastie2009elements}	& GBT & SciKit & 11 \\
		Extremely randomised trees~\cite{geurts2006extremely}	& ERT & SciKit & 7 \\
		AdaBoosted decision trees$^\text{a}$~\cite{drucker1997improving}	& ABT & SciKit & 3 \\
		Gaussian process regression~\cite{williams2006gaussian}	& GPR & SciKit & 2 \\
		$k$ nearest neighbours	& KNN & SciKit & 3 \\
		Artificial neural networks	& ANN & Keras~\cite{chollet2015keras} & 2 \\
		Inverse distance weighing~\cite{shepard1968two} & IDW & SMT~\cite{SMT2019} & 1 \\
		Radial basis functions & RBF & SMT & 3 \\
		\bottomrule
		\end{tabular}\\%
		{\footnotesize $^\text{a}$Note that ABTs can be viewed as a subclass of GBTs.}
	\end{indented}
\end{table}

@@

\begin{table*}[t]
	\renewcommand{\arraystretch}{0.95}
	\caption{\label{tbl:metrics}Metrics recorded in experiments. In
	formulations, we work with a training set of size $N_0$ and a test set of
size $N$, values $y^{(i)}=f(x^{(i)})$ and $\hat{y}^{(i)}=\hat{f}(x^{(i)})$
denote images of the $i$th testing sample in the MC TBR model and the surrogate
respectively. The mean $\overline{y}=N^{-1}\sum_{i=1}^N y^{(i)}$ and $P$ is the
number of input features.}
	\begin{indented}
	\item[]
		\begin{tabularx}{\textwidth}{Xrl}
		\toprule
		Regression performance metrics& Notation	& Mathematical formulation\\
		\midrule
		Mean absolute error	& MAE & $N^{-1}\sum_{i=1}^N |y^{(i)}-\hat{y}^{(i)}|$ \\
		Standard deviation of error & $S$	& $\text{StdDev}_{i=1}^N\left\{ |y^{(i)} -
		\hat{y}^{(i)}| \right\} $ \\
			Coefficient of determination & $R^2$	& $1-\sum_{i=1}^N
			\left(y^{(i)}-\hat{y}^{(i)} \right)^2\left[\sum_{i=1}^N \left(
			y^{(i)}-\overline{y} \right)^2\right]^{-1} $ \\
			Adjusted $R^2$ & $R^2_\text{adj.}$	& $1-(1-R^2)(N-1)(N-P-1)^{-1}$ \\
		\midrule
		Computational complexity metrics	& {}	& {} \\
		\midrule
		Mean training time & $\overline{t}_{\text{trn.}}$	& $(\text{wall training time of
		$\hat{f}(x)$})N_0^{-1}$  \\
			Mean prediction time & $\overline{t}_{\text{pred.}}$	& $(\text{wall prediction time of
			$\hat{f}(x)$})N^{-1}$ \\
				Relative speedup & $\omega$	& $(\text{wall evaluation$^\text{b}$ time of $f(x)$})
				(N\overline{t}_{\text{pred.}})^{-1}$ \\
		\bottomrule
		\end{tabularx}\\%
		{\footnotesize $^\text{b}$This corresponds to evaluation of the
		expensive MC TBR model on all points of the test set. In surrogates, the
		same time period is referred to as ``prediction time.''}
	\end{indented}
\end{table*}

@@

The objective of Experiment~1 is to simplify the regression task for
surrogates prone to suboptimal performance in discontinuous spaces.
To this end, training points are filtered to a single selected discrete feature
assignment, and surrogates are trained only on the remaining continuous features.
This is repeated several times to explore variances in behaviour,
particularly in 4~distinct assignments that are obtained by setting blanket and
firstwall coolant materials to one
of:~$\{\text{H\textsubscript{2}O},\isotope{He}\}$.
Experiment~2 conventionally measures surrogate performance on the full feature
space without any parameter restrictions. In both experiments, hyperparameter tuning is
facilitated by Bayesian optimisation~\cite{movckus1975bayesian}. We select the
hyperparameter configuration that produces the model that maximises $R^2$. The
process is terminated after 1000~iterations or 2~days, whichever condition is satisfied first.
The results of Experiments~1 \&~2 are depicted
in Figures~\ref{fig:exp1-time-vs-reg} \&~\ref{fig:exp2-time-vs-reg}
respectively, and described in~\Sref{sec:res-exp12}.

In Experiment~3, the 20~best-performing hyperparameter configurations
per each model family are used to train surrogates on sets of various sizes to
investigate their scaling properties. In particular, we consider sets of
sizes~$\{1,2,5,10,12,15,20\}$ in thousands of samples to better characterize the
relationship of each family between sample count and the metrics of interest
listed in~\Tref{tbl:metrics}.
The results are shown in~\Fref{fig:scaling} and discussed
in~\Sref{sec:res-exp3}.

Finally, Experiment~4 aims
to produce surrogates suitable for practical use by retraining selected
well-scaling instances on large training sets in order to satisfy the goals of this work.
The results are displayed in~\Fref{fig:reg-performance} and
in~\Tref{tbl:exp4-detailed-results}, and summarized in~\Sref{sec:res-exp4}.
